Intro Statistical Analysis Script

### 2024-11-14 code/analysis-code/fullanalysis-v1.qmd
RRSSM2
Age2; immigrants; year fixed effects

$$\text{BW} = \alpha + \beta_1 \text{MA} + \beta_2 \text{Gest} + \beta_3 \text{Nat} + \beta_4 \text{Age} + \epsilon$$

whereas `BW` is birth weight in grams, $\alpha$ is the average birth weight, $\beta_x$ are the group deviations from the average birth weight, `MA` is maternal ancestry (Euro-descent, Afro-descent), `Gest` is gestational status (nulliparous or multiparous), `Age` is maternal age in years, and $\epsilon$ is the error term. Results are displayed in Table 3.

@@

# Using recipes and workflows
# Define recipe for model
ML_recipe2 <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(ModifiedColor) #%>% # Convert all nominal variables to dummy variables. Cannot do this if I am merging tables in gtsummary
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors. Cannot do this if I am merging tables in gtsummary

@@

# Using recipes and workflows
# Define recipe for model
ML_recipe2_train <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm2_sub <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ ModifiedColor + Age + + ModifiedStatus, data = filtered_data)

@@

The second two models are also simple and multiple linear regressions with outcome variable as birthweight in grams. However, in these two models, I maintained maternal skin color as three categories, White (the reference group), Black, and mixed race to see if there is differential outcomes for Black and mixed-race women. The first simple linear regression model is as follows:

@@

$$\text{BW} = \alpha + \beta_1 \text{MC} + \beta_2 \text{Gest} + \beta_3 \text{Nat} + \beta_4 \text{Age} + \epsilon$$

whereas $\alpha$ is the average birth weight, $\beta_x$ are the group deviations from the average birth weight, `BW` is birth weight in grams, `MC` is maternal skin color (White, Black, and mixed race), `Gest` is gestational status (nulliparous or multiparous), `Nat` is maternal nationality (Brazilian, Latin American, European, Middle Eastern), `Age` is maternal age in years, and $\epsilon$ is the error term.

@@

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4 <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") #%>% # Set reference level for ModifiedStatus
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_train <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Define recipe for model
ML_recipe4_cv <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_test <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = test_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

>>>>

$$\text{BW} = \alpha + \beta_1 \text{MA} + \beta_2 \text{Par} + \beta_3 \text{Nat} + \beta_4 \text{Age} + \text{YrFE} + \epsilon$$

whereas `BW` is birth weight in grams, $\alpha$ is the average birth weight, $\beta_x$ are the group deviations from the average birth weight, `MA` is maternal ancestry (Euro-descent, Afro-descent), `Par` is parity (nulliparous or multiparous), `Age` is maternal age in years, `YrFE` is year fixed effect, and $\epsilon$ is the error term. Results are displayed in Table 3.

@@

# Using recipes and workflows
# Define recipe for model
ML_recipe2 <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality + Year, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(ModifiedColor) #%>% # Convert all nominal variables to dummy variables. Cannot do this if I am merging tables in gtsummary
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors. Cannot do this if I am merging tables in gtsummary

@@

# Using recipes and workflows
# Define recipe for model
ML_recipe2_train <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality + Year, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm2_sub <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + Year, data = filtered_data)

@@

The second two models are also simple and multiple linear regressions with outcome variable as birth weight in grams. However, in these two models, I maintained maternal skin color as three categories, White (the reference group), Black, and mixed race to see if there is differential outcomes for Black and mixed-race women. The first simple linear regression model is as follows:

@@

$$\text{BW} = \alpha + \beta_1 \text{MC} + \beta_2 \text{Par} + \beta_3 \text{Nat} + \beta_4 \text{Age} + \text{YrFE} + \epsilon$$

whereas $\alpha$ is the average birth weight, $\beta_x$ are the group deviations from the average birth weight, `BW` is birth weight in grams, `MC` is maternal skin color (White, Black, and mixed race), `Par` is parity (nulliparous or multiparous), `Nat` is maternal nationality (Brazilian, Latin American, European, Middle Eastern), `Age` is maternal age in years, `YrFE` is year fixed effect, and $\epsilon$ is the error term.

@@

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4 <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality + Year, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") #%>% # Set reference level for ModifiedStatus
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_train <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality + Year, data = train_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") #%>% # Set reference level for ModifiedStatus
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Define recipe for model
ML_recipe4_cv <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality + Year, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") #%>% # Set reference level for ModifiedStatus
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_test <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality + Year, data = test_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") #%>% # Set reference level for ModifiedStatus
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

### 2024-11-14 code/analysis-code/introanalysis-v1.qmd
RRSSM2
Age2; immigrants; year fixed effects

```{r}
# See LBW and NBW with true denominator
LBW_percentage <- ML_summary %>%
  filter(!is.na(ModifiedColor) & !is.na(BirthweightCategory)) %>% # Remove NAs from both variables
group_by(ModifiedColor) %>% # Group by ModifiedColor

  # Calculate the percentage for each group
summarise(
  total_count = n(),
  lbw_count = sum(BirthweightCategory == "LBW"),
  lbw_percentage = (lbw_count / total_count) * 100
  )

@@

# Linear Models

@@

@@

```{r}
# Model 1 using recipes and workflows
# Define recipe for model
ML_recipe1 <- recipe(Weightgrams ~ ModifiedColor, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

@@

# Save regression output into table
saveRDS(table_fit1, file = here::here("results", "tables", "table_fit1.rds"))
```

@@

```{r}
# Model Age using recipes and workflows
# Create recipe
ML_recipe_age <- recipe(Weightgrams ~ Age, data = ML_linear)

@@

# Save regression output into table
saveRDS(table_fit2, file = here::here("results", "tables", "table_fit2.rds"))
```

Now I will run a slightly more complicated model that includes maternal skin color, age, nationality, and gestational status.

```{r}
# Set reference level for ModifiedColor
ML_linear$ModifiedColor <- relevel(ML_linear$ModifiedColor, ref = "Euro-Descent")

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm3 <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear)

@@

In the first and third models, there appears to be an association between maternal skin color and infant birth weight, with Euro-descended women (the reference group) having infants with higher birth weights than Afro-descended women. In the second model, we can see that older mothers are associated with giving birth to infants with higher birth weights.

@@

@@

These results imply that Europeans had heavier infants than Brazilians. Since all women of color in this sample were Brazilian, there is a problem with multicollinearity. So I will perform some sub-sample modeling to understand this issue. We will create a `Brazilian` (yes, no) variable, drop all non-Brazilians from the study, and run the original regression with the covariates.

```{r}
# Create dummy variables for new category of Brazilian or not
# Create dummy variables for status and change to English
ML_linear <- mutate(ML_linear, 
                    Brazilian = recode_factor(ModifiedNationality, 
                                              "European" = "Non-Brazilian",
                                              "Middle Eastern" = "Non-Brazilian",
                                              "Latin American" = "Non-Brazilian",
                                              "Brazilian" = "Brazilian"))

# Check the structure of the updated data frame
str(ML_linear)

# Save ML_Linear as RDS
save_data_location <- here::here("data","processed-data","ML_linear.rds")
saveRDS(ML_linear, file = save_data_location)
```

@@

```{r}
# Model 5 sub-sample simple model
# Set reference level for Brazilian
ML_linear$Brazilian <- relevel(ML_linear$Brazilian, ref = "Brazilian")

@@

@@

```{r}
# Filter data based on desired conditions
filtered_data <- ML_linear %>%
  filter(Brazilian == "Brazilian")

@@

## Multivariate analysis

@@

@@

# Save regression output into table
saveRDS(table_fit7, file = here::here("results", "tables", "table_fit7.rds"))

# Complex linear model
# Set reference level for Color
ML_linear$Color <- relevel(ML_linear$Color, ref = "White")

@@

## Logistic models

@@

@@

@@

# Store metrics in a list for dynamic inline text
logistic_model_metrics2 <- list(
  pseudo_r_squared = mcfadden_r2_1,
  auc = roc_auc_result1$.estimate
)
all_logistic_models$model2 <- logistic_model_metrics2 # add to list of lists
```

@@

@@

# Store metrics in a list for dynamic inline text
logistic_model_metrics3 <- list(
  pseudo_r_squared = mcfadden_r2_1,
  auc = roc_auc_result1$.estimate
)
all_logistic_models$model3 <- logistic_model_metrics3 # add to list of lists
```

@@

@@

# Create recipe
# Create recipe simple logistic model (creating dummy variables and standardizing continuous variables)
log4_recipe <- recipe(BirthweightCategory ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_normalize(all_predictors())

@@

# Store metrics in a list for dynamic inline text
logistic_model_metrics4 <- list(
  pseudo_r_squared = mcfadden_r2_1,
  auc = roc_auc_result1$.estimate
)
all_logistic_models$model4 <- logistic_model_metrics4 # add to list of lists
```

@@

# Merge unadjusted log models 1 and 2

@@

>>>>

```{r}
# See LBW and NBW with true denominator
LBW_percentage <- ML_summary %>%
  filter(!is.na(ModifiedColor) & !is.na(BirthweightCategory)) %>% # Remove NAs from both variables
  group_by(ModifiedColor) %>% # Group by ModifiedColor
  summarise( # Calculate the percentage for each group
    total_count = n(),
    lbw_count = sum(BirthweightCategory == "LBW"),
    lbw_percentage = (lbw_count / total_count) * 100
    )

@@

# Linear Models ModifiedColor

@@

## Model 1

@@

```{r}
# Model 1 using recipes and workflows
# Define recipe for model
ML_recipe1 <- recipe(Weightgrams ~ ModifiedColor, data = ML_linear) #%>% 
  #step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level
  #step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

@@

# Save regression output into table
saveRDS(table_fit1, file = here::here("results", "tables", "table_fit1.rds"))
```
## Model Age

@@

```{r}
# Model Age using recipes and workflows
# Create recipe
ML_recipe_age <- recipe(Weightgrams ~ Age + AgeSquared, data = ML_linear)

@@

# Save regression output into table
saveRDS(table_fit2, file = here::here("results", "tables", "table_fit2.rds"))
```
## Model 3

Now I will run a slightly more complicated model that includes maternal skin color, age, age squared, nationality, and parity (`ModifiedStatus`). This is the basis for my main model.

```{r}
# Model 3 using recipes and workflows
# Set reference level for ModifiedColor
ML_linear$ModifiedColor <- relevel(ML_linear$ModifiedColor, ref = "Euro-Descent")

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm3 <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ ModifiedColor + Age + AgeSquared + ModifiedStatus + ModifiedNationality, data = ML_linear)

@@

In the first and third models, there appears to be an association between maternal skin color and infant birth weight, with Euro-descended women (the reference group) having infants with higher birth weights than Afro-descended women. In the second model, we can see that older mothers are associated with giving birth to infants with higher birth weights. In the third model, neither age nor age-squared are significant. Nonetheless, birthweight appears to increases with age, but above the maternal age of 40, infant weight decreases. Thus, there is a reverse U-shaped relationship. The coefficients suggests that while birthweight increases with maternal age, it does so at a decreasing rate until a certain maternal age after which it begins decreasing.

@@

### Model 4

@@

These results imply that Europeans had heavier infants than Brazilians. Since all women of color in this sample were Brazilian, there is a problem with multicollinearity. So I will perform some sub-sample modeling to understand this issue. In `eda-v1.qmd` I created a `Brazilian` (yes, no) variable, so here, I will drop all non-Brazilians from the study, and run the original regression with the covariates.

### Model 5

@@

```{r}
# Model 5 sub-sample simple model
# Set reference level for Brazilian
ML_linear$Brazilian <- relevel(ML_linear$Brazilian, ref = "yes")

@@

### Model 6

@@

```{r}
# Filter data based on desired conditions
filtered_data <- ML_linear %>%
  filter(Brazilian == "yes")

@@

# Linear Models Color

@@

## Model 7

@@

# Save regression output into table
saveRDS(table_fit7, file = here::here("results", "tables", "table_fit7.rds"))
```

## Model 8

```{r}
# Model 8 multivariate linear model
# Set reference level for Color
ML_linear$Color <- relevel(ML_linear$Color, ref = "White")

@@

# Year Fixed-Effect

## Model Years (9, 10, 11)

In the first round of peer review, one reviewer suggested that I should justify why I provide this as a cross-sectional approach rather than test for fixed effects by year given periodic epidemics in infectious disease such as malaria and yellow fever and changes in real wages. So I will run a model that includes year as a fixed effect here with the `ModifiedColor` variable as the main predictor variable.

```{r}
# Model 9: multivariate linear model with year as fixed effect using ModifiedColor
ML_linear <- ML_linear %>%
  mutate(Year = as.factor( # Create a new column called "Year" and make it a factor
    substr(Date, 1, 4) # Take the first 4 characters of the Date column
  ))

# Save new variable into dataset
save_data_location <- here::here("data","processed-data","ML_linear.rds")
saveRDS(ML_linear, file = save_data_location)

# Set reference level for Color
ML_linear$ModifiedColor <- relevel(ML_linear$ModifiedColor, ref = "Euro-Descent")

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm9 <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality + Year, data = ML_linear)

# Summarize model fit with tidiers
ML_lm9_summary <- summary(ML_lm9) # Summary of fit
tidy(ML_lm9, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
model_stats9 <- glance(ML_lm9) # Extract model-level statistics

# Extract R-squared value
r_squared_values3$model9 <- model_stats9$r.squared

# Create a gtsummary table directly from the model
table_fit9 <- tbl_regression(ML_lm9, exponentiate = FALSE, intercept = TRUE) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error))
table_fit9

# Save regression output into table
saveRDS(table_fit9, file = here::here("results", "tables", "table_fit9.rds"))
```

I will also run the model five times, once per year.

```{r}
# Model 10: multivariate linear model for each year (1922-1924) with ModifiedColor
# Set reference level for Color
ML_linear$ModifiedColor <- relevel(ML_linear$ModifiedColor, ref = "Euro-Descent")

# Ensure year is created as character vector
years <- c("1922", "1923", "1924", "1925", "1926")

# Create models for all years at once
table_years <- ML_linear %>%
  mutate(Year = as.character(substr(Date, 1, 4))) %>% # Create a new column called "Year" and make it a factor
  filter(Year %in% years) %>%
  group_by(Year) %>%
  nest() %>%
  mutate(
    model = map(data, ~linear_reg() %>% 
                  set_engine("lm") %>% 
                  fit(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, 
                      data = .x)),
    tables = map(model, ~tbl_regression(.x, 
                                      exponentiate = FALSE, 
                                      intercept = TRUE) %>% 
                   bold_p() %>% 
                   add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
                   modify_column_unhide(columns = c(statistic, std.error)))
  )

# To view results for a specific year:
table_years %>% 
  filter(Year == "1922") %>% 
  pull(tables) %>% 
  .[[1]]

# Create named list
tables_list <- table_years %>%
  arrange(Year) %>%
  pull(tables) %>%
  set_names(paste0("Year_", years))

# Create gt summary table
combined_table_years <- tbl_merge(
 tbls = tables_list,
 tab_spanner = c("**1922**", "**1923**", "**1924**", "**1925**", "**1926**")
) %>%
 modify_table_styling(
   columns = everything(),
   align = "center"
 ) %>%
 modify_table_body(fun = ~.x %>% arrange(variable))
combined_table_years

# Save regression output into table
saveRDS(combined_table_years, file = here::here("results", "tables", "combined_table_years.rds"))
```

Given that 1923 demonstrates a coefficient for infant birth weights to Afro-descended mothers nearly 3 times the overall coefficient (-84 to -237), I will drop the observations from 1923 and re-run the regression to see if the results are still robust. The coefficient for infants born to Afro-descended mothers is not significant and is lower than in the full model (-40). This demonstrates that the births from 1923 seem to be driving the overall results. Nonetheless, the overall hypothesis still holds. Infants born to Afro-descended mothers had lower birth weights than those born to Euro-descended mothers, both in years where all birth weights were higher and years in which all birth weights were lower. Economic uncertainty affected women of African descent more than those of European descent. Additionally, these results could be due to smaller sample sizes and larger standard errors -- noise in the data.

```{r}
# Model 11: Multivariate model using ModifiedColor without observations from 1923
# Filter out observations from 1923
ML_linear_filtered <- ML_linear %>%
  filter(Year != "1923")

# Set reference level for Color
ML_linear$ModifiedColor <- relevel(ML_linear$ModifiedColor, ref = "Euro-Descent")

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm11 <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality + Year, data = ML_linear_filtered)

# Summarize model fit with tidiers
ML_lm11_summary <- summary(ML_lm11) # Summary of fit
tidy(ML_lm11, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
model_stats11 <- glance(ML_lm11) # Extract model-level statistics

# Extract R-squared value
r_squared_values3$model11 <- model_stats11$r.squared

# Create a gtsummary table directly from the model
table_fit11 <- tbl_regression(ML_lm11, exponentiate = FALSE, intercept = TRUE) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error))
table_fit11

# Save regression output into table
saveRDS(table_fit11, file = here::here("results", "tables", "table_fit11.rds"))
```

# Logistic models

@@

## Log Model 1

@@

## Log Model 2

@@

# Store metrics in a list for dynamic inline text
logistic_model_metrics2 <- list(
  pseudo_r_squared = mcfadden_r2_2,
  auc = roc_auc_result1$.estimate
)
all_logistic_models$model2 <- logistic_model_metrics2 # add to list of lists
```

@@

## Log Model 3

@@

# Store metrics in a list for dynamic inline text
logistic_model_metrics3 <- list(
  pseudo_r_squared = mcfadden_r2_3,
  auc = roc_auc_result1$.estimate
)
all_logistic_models$model3 <- logistic_model_metrics3 # add to list of lists
```

@@

## Log Model 4

@@

# Create recipe
# Create recipe simple logistic model (creating dummy variables and standardizing continuous variables)
log4_recipe <- recipe(BirthweightCategory ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) #%>%
 # step_dummy(all_nominal(), -all_outcomes()) %>%
 # step_normalize(all_predictors())

@@

# Store metrics in a list for dynamic inline text
logistic_model_metrics4 <- list(
  pseudo_r_squared = mcfadden_r2_4,
  auc = roc_auc_result1$.estimate
)
all_logistic_models$model4 <- logistic_model_metrics4 # add to list of lists
```

@@

## Merge unadjusted log models 1 and 2

@@

## Log Model Stilbirth

Reviewer 3 wrote: "A logit analysis using stillbirth as an outcome would bring robustness to the racial disparities that the author highlights. That there are 221 stillbirths creates this possibility and might even reveal greater racial disparities. While these could not be definitively linked to syphilis, this might provide further insight into that data limitation. Again, causation cannot be overstated, but given that the author cannot speak to prematurity with the data (another important indicator of maternal stress), this would be an interesting avenue to explore."

```{r}
# Check factor levels
levels(ML_summary$ModifiedColor) # Reference is Euro-Descent
levels(ML_summary$FetalOutcome) # Reference is Live Birth

# Set the reference level explicitly to NBW
ML_summary_log <- ML_summary %>%
  filter(!is.na(FetalOutcome) & !is.na(ModifiedColor)) %>% #filter out NAs
  mutate(FetalOutcome = relevel(FetalOutcome, ref = "Live Birth"),
         ModifiedColor = relevel(ModifiedColor, ref = "Euro-Descent"))

#set seed
set.seed(456)
ML_log_split2 <- initial_split(ML_summary_log, prop = 0.75, strata = FetalOutcome)

# Create training and test datasets
ML_log_train2 <- training(ML_log_split2)
ML_log_test2 <- testing(ML_log_split2)

# Logistic model 5: simple logistic regression model with stillbirth as outcome and ModifiedAncestry as the predictor
# Create recipe
# Create recipe simple logistic model (creating dummy variables and standardizing continuous variables)
log5_recipe <- recipe(FetalOutcome ~ ModifiedColor, data = ML_summary_log) # %>%
  #step_dummy(all_nominal(), -all_outcomes()) %>%
  #step_normalize(all_predictors())

# Define model specification
log5_spec <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

# Create workflow
log5_wf <- workflow() %>%
  add_recipe(log5_recipe) %>%
  add_model(log5_spec)

# Fit model
log5_fit <- log5_wf %>%
fit(data = ML_summary_log)

# Extract fitted model from workflow
log5_fitted <- extract_fit_parsnip(log5_fit)$fit

# Create a gtsummary table directly from the model
table_log5 <- tbl_regression(log5_fitted, exponentiate = TRUE) # exponentiate true for ORs

#Bold significant p-values and add n
table_log5 <- bold_p(table_log5, t = 0.05) %>%
  add_n()
table_log5

# Save regression output into table
saveRDS(table_log5, file = here::here("results", "tables", "table_log5.rds"))

# Calculate metrics using pseudo-R2
# Extract model metrics using broom
model_metrics5 <- glance(log5_fitted)

# Calculate McFadden's R-squared
# McFadden's R-squared = 1 - (deviance(model) / deviance(null model))
mcfadden_r2_5 <- 1 - (model_metrics5$deviance / model_metrics5$null.deviance)

# Print McFadden's R-squared
print(mcfadden_r2_5)

# Calculate performance metrics using ROC_AUC
# Fit the model on the training data
log5_fit <- log5_wf %>% fit(data = ML_log_train2)

# Predict probabilities on the test set
prob_preds5 <- predict(log5_fit, new_data = ML_log_test2, type = "prob")

# Combine predictions with true values
results5 <- bind_cols(ML_log_test2, prob_preds5)

# Calculate ROC-AUC
roc_auc_result5 <- roc_auc(results5, truth = FetalOutcome, `.pred_Live Birth`)

# Print ROC-AUC
print(roc_auc_result5)

# Create and plot the ROC curve
roc_curve_data5 <- roc_curve(results5, truth = FetalOutcome, `.pred_Live Birth`)

# Plotting the ROC curve
roc_plot5 <- ggplot(roc_curve_data5, aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = 2) +
  labs(title = "ROC Curve",
       x = "1 - Specificity",
       y = "Sensitivity")

# Print the plot
print(roc_plot5)

# Store metrics in a list for dynamic inline text
logistic_model_metrics5 <- list(
  pseudo_r_squared = mcfadden_r2_5,
  auc = roc_auc_result1$.estimate
)
all_logistic_models$model5 <- logistic_model_metrics5 # add to list of lists1
```
